---
title: "Homework_06"
author: "Jasmine Hart"
date: "2/23/2022"
output: html_document
---

# 1.

```{r, echo=TRUE}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

# Step 2 - Random data set example *
```{r, echo=TRUE}
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

# Plot histogram of data
```{r z, echo=TRUE}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```
# add empirical density curve
```{r, echo=TRUE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

# maximum likelihood paramaeters for "normal"
```{r, echo=TRUE}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute ex. mean
```

# Plot normal probability denisty
```{r, echo=TRUE}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```
 
 # Plot exponential probability density
```{r, echo=TRUE}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```
 
 # Plot uniform prob density
```{r, echo=TRUE}
 stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```
 
# Plot gamma probability denisty
```{r, echo=TRUE}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```
 
 # Plot beta probability density 
```{r, echo=TRUE}
 pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

 --------------- Own data set -----------------
Data set from dryad. The paper is "Repeated genetic and adaptive phenotypic divergence across tidal elevation in a foundation plant species" I selected the maternal plant height at the time of seed collection. They used the offspring to evaluate height and other characteristics to better understand genetic divergence of this marsh species.

# 2.

```
z <- read.table("Maternal_plants_field.csv",header=TRUE,sep=",")
str(z)
summary(z)
```

```{r, echo=TRUE}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

# Step 2 - Random data set example *

```{r, echo=TRUE}
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

# Plot histogram of data
```{r, echo=TRUE}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

# add empirical density curve
```{r, echo=TRUE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

# maximum likelihood paramaeters for "normal"
```{r, echo=TRUE}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute ex. mean
# giving mean of maternal plant height 86.6 cm
```

# Plot normal probability denisty
```{r, echo=TRUE}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat
```
 
# Plot exponential probability density
```{r, echo=TRUE}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2
```
 
 # Plot uniform prob density
```{r, echo=TRUE}
 tat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```
 
# Plot gamma probability denisty
```{r, echo=TRUE}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```
 
 # Plot beta probability density 
```{r, echo=TRUE}
 pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

# 3.

```{r, echo=TRUE}
# the gamma probability density fits the data set the best 

gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

# 4.

```{r, echo=TRUE}
# quick and dirty, a truncated normal distribution to work on the solution set

y <- rnorm(n=1732,mean=0.8661687,0.62406564)
y <- data.frame(1:1732,y)
names(y) <- list("ID","myVar")
y <- y[y$myVar>0,]
str(y)
summary(y$myVar)
```

# probability denisty curve
```{r, echo=TRUE}
gammaPars1 <- fitdistr(y$myVar,"gamma")
shapeML1 <- gammaPars1$estimate["shape"]
rateML1 <- gammaPars1$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(y$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```
 
 
# The two graphs look very similar, but the original data set has a bit more variation and the simulated dated looks a lot more uniform than the other. Which makes sense because it is a randomly generated integer sequence.